---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fleeteffSim

<!-- badges: start -->
[![R-CMD-check](https://github.com/grattan/fleeteffSim/workflows/R-CMD-check/badge.svg)](https://github.com/grattan/fleeteffSim/actions)
<!-- badges: end -->

Things to be fixed:

 - add a test so that the select_upgrade function always returns the absolute lowest value (
 particularly important as I have added 'shortcuts' to getting to the lowest cost which may
 not always hold if future changes are made so that vehicles in the same type have different base emissions)
 
 - need to make sure that there is a specific and known protocol for how the algorithm decides 
 which car to apply the electric upgrade to when they are all the same cost (i.e. all 0 cost)
 
 - need to add rebound effect
 
 - need to add a 1 year earlier and 1 year later price parity scenario to ev curves




## Installation

You can install the released version of `fleeteffSim` from [Github]() with:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("grattan/fleeteffSim")
```

## Example

The `fleet_eff_sim()` function runs simulations of vehicle emissions standards over the Australian fleet. By default, this uses a grattan trajectory listed in the 'Grattan car plan' report as the 'central' scenario, and simulates the use of vehicles sold between 2024 and 2050. 

```{r example, message = FALSE}

library(fleeteffSim)
## basic example code

test_run <- fleet_eff_sim()
```

The resulting `tibble` is:

``` {r glimpse}

glimpse(test_run)

```

There are many inputs to the `fleet_eff_sim()` function that can be changed from default settings. For example, this includes future fuel and electricity prices, as well as forecast future sales volumes by vehicle category. 


The `fleet_eff_sim()` function only provides summarised data over the entire period run in the model. 
For more granular output, the following series of functions can be run. 

First, production costs required to meet targets can be estimated:

``` {r compliance}

#Running the target scenario to estimate vehicle production costs
target_compliance <- compliance_costs(fleet = fleet_creator(),
                                      .target_scenario = "target_central")

#Running the 'business as usual'  scenario to estimate vehicle production costs
bau_compliance <- compliance_costs(fleet = fleet_creator(),
                                   .target_scenario = "bau")

```


Second, running costs and emissions can be estimated:

``` {r compliance}

#Running the target scenario to estimate vehicle production costs
target_running <- benefit_model(fleet = target_compliance)

#Running the 'business as usual'  scenario to estimate vehicle production costs
bau_running <- benefit_model(fleet = bau_compliance)

```


The `target_running` and `bau_running` datasets can then be compared for more granular data - however, it must be considered that each 'simulated' vehicle under default parameters accounts for 1\% of the total estimated fleet. Thus, runnings costs and emissions must be scaled up respectively to estimated fleet wide values. 

Similarly to the `fleet_eff_sim()` function, summarised results can also be generated. This accounts for scaling up the fleet:

``` {r summary}

generate_results(bau_benefits = bau_running,
                 target_benefits = target_running,
                 cars = 100)

```

